The low-frequency Murchison radio environment:\\interference analysis and mitigation

### 2014-10-03 articles/2014/MWA-Radio-Environment/mwa-radio-environment.tex
Article

\author[A.~R.~Offringa et al.]{A.~R.~Offringa$^{1,2}$\thanks{\editmark{Corresponding author.} E-mail: \url{andre.offringa@anu.edu.au}},
R.~B.~Wayth\Curtin$^,$\CAASTRO, ... , %\newauthor
builders list
\\
\ANU{}Research School of Astronomy and Astrophysics, Australian National University, Canberra, ACT 2611, Australia \\
\CAASTRO{}ARC Centre of Excellence for All-sky Astrophysics (CAASTRO), Australian National University, Canberra, ACT 2611, Australia \\
\Curtin{}International Centre for Radio Astronomy Research, Curtin University, Bentley, WA 6102, Australia\\
}

@@

\section{Method}
RFI detection, often referred to as ``data flagging'', is one of the first steps in the data processing. It is important to perform initial data flagging at high time and frequency resolution, because this increases detection accuracy and decreases the loss of data \citep{lofar-radio-environment}. Consequently, RFI detection has to work on large data volumes, and its computational cost is therefore a concern -- in particular for a many-element arrays. In some projects, simple thresholding is used to mitigate the worst interference, which is computationally cheap but not very accurate. For example, a $3\sigma$ threshold is used for reducing PAPER data in \citet{parsons-paper-eorlimit-2014}. Several observatories or projects have designed pipelines that include more advanced RFI mitigation. Examples of such pipelines include \textsc{aoflagger} \citep{post-correlation-rfi-classification,scale-invariant-rank-operator}, originally designed for LOFAR; \textsc{flagcal} for preprocessing data from the Giant Metrewave Radio Telescope (GMRT; \citealt{prasad-flagcal-2012}); and \textsc{serpent} for preprocessing data from the Multi-Element Radio Linked Interferometer Network (e-MERLIN; \citealt{serpent-peck-2013}). For RFI detection in MWA observations, LOFAR's \textsc{aoflagger} is used. This flagger has shown to have a good accuracy, is fast and has a library interface, which allows it to be integrated in a pipeline.

@@

\begin{table*}
\caption{List of observations used in the analyses. The RFI column contains the fraction of visibilities that the initial \textsc{aoflagger} run has classified as RFI (which includes false positives). The DVB column displays the fraction of visibilities that were unusable because of the presence of DVB signals, and which the flagger does not flag. Because the DVB frequencies are only observed in GLEAM observations, this fraction is only given for those. }\label{tbl:obs-list}
\begin{tabular}{|l|l|r|r|r|c|}
\hline
\textbf{Project} &\textbf{Date} & \textbf{Frequencies (MHz)} & \textbf{Duration} & RFI & DVB \\
\hline
EoR   & 2013-08-23 & 138.8--197.7              & 3h & 0.53\% & ---\\
GLEAM & 2013-08-25 & 72.3--133.8, 138.8--230.8 & 7h & 0.94\% & 0\% \\
EoR   & 2013-08-26 & 138.8--197.7              & 6h & 0.81\% & ---\\
GLEAM & 2013-11-05 & 72.3--133.8, 138.8--230.8 & 7h & 1.27\% & 0\% \\
GLEAM & 2013-11-25 & 72.3--133.8, 138.8--230.8 & 7h & 0.69\% & 0\% \\
EoR   & 2014-02-05 & 138.8--197.7              & 6h & 0.54\% & -- \\
GLEAM & 2014-03-16 & 72.3--133.8, 138.8--230.8 & 7h & 0.79\% & 0\% \\
GLEAM & 2014-03-17 & 72.3--133.8, 138.8--230.8 & 7h & 1.64\% & 3\% \\
EoR   & 2014-04-10 & 138.8--197.7              & 6h & 0.68\% & ---\\
GLEAM & 2014-06-18 & 72.3--133.8, 138.8--230.8 & 7h & 0.98\% & 0\% \\
RFI test&2014-08-27& 131.2--161.9     & 2$\times$8 m& 1.85\% & --- \\
\hline
\multicolumn{3}{|r|}{\textbf{Total}} & \textbf{63h} & \textbf{...\%} & \textbf{0.5\%}\\
\hline
\end{tabular}
\end{table*}

@@

\noindent\begin{figure*}
\begin{center}\hspace*{-0.2cm}\includegraphics[width=18cm]{img/plot-rfi-per-set-eor}\vspace{-1cm}
\caption{..}
\label{fig:rfi-per-set-eor}
\end{center}
\end{figure*}

\noindent\begin{figure*}
\begin{center}\hspace*{-0.2cm}\includegraphics[width=18cm]{img/plot-rfi-per-set-gleam}\vspace{-1cm}
\caption{..}
\label{fig:rfi-per-set-gleam}
\end{center}
\end{figure*}

@@

\noindent\begin{figure}
\begin{center}\hspace*{-0.2cm}\includegraphics[width=9.1cm]{img/rfi-per-band}
\caption{RFI per subband}
\label{fig:rfi-per-band}
\end{center}
\end{figure}

\noindent\begin{figure}
\begin{center}\hspace*{-0.2cm}\includegraphics[width=9.1cm]{img/lognlogs-residual}
\caption{Log N log S for residuals}
\label{fig:lognlogs-residuals}
\end{center}
\end{figure}

\noindent\begin{figure}
\begin{center}\hspace*{-0.2cm}\includegraphics[width=9.1cm]{img/lognlogs-rfi}
\caption{Log N log S for RFI }
\label{fig:lognlogs-rfi}
\end{center}
\end{figure}

>>>>

\author[A.~R.~Offringa et al.]{A.~R.~Offringa$^{1,2}$\thanks{\editmark{Corresponding author.} E-mail: \url{andre.offringa@anu.edu.au}},
R.~B.~Wayth\Curtin$^,$\CAASTRO,
N.~Hurley-Walker\Curtin,
D.~Kaplan\Curtin,
... , %\newauthor
builders list
\\
\ANU{}Research School of Astronomy and Astrophysics, Australian National University, Canberra, ACT 2611, Australia \\
\CAASTRO{}ARC Centre of Excellence for All-sky Astrophysics (CAASTRO), Australian National University, Canberra, ACT 2611, Australia \\
\Curtin{}International Centre for Radio Astronomy Research, Curtin University, Bentley, WA 6102, Australia\\
}

@@

\section{Method}
RFI detection, often referred to as ``data flagging'', is one of the first steps in the data processing. It is important to perform initial data flagging at high time and frequency resolution, because this increases detection accuracy and decreases the loss of data \citep{lofar-radio-environment}. Consequently, RFI detection has to work on large data volumes, and its computational cost is therefore a concern -- in particular for many-element arrays. In some projects, simple thresholding is used to mitigate the worst interference, which is computationally cheap but not very accurate. For example, a $3\sigma$ threshold is used for reducing PAPER data in \citet{parsons-paper-eorlimit-2014}. Several observatories or projects have designed pipelines that include more advanced RFI mitigation. Examples of such pipelines include \textsc{aoflagger} \citep{post-correlation-rfi-classification,scale-invariant-rank-operator}, originally designed for LOFAR; \textsc{flagcal} for preprocessing data from the Giant Metrewave Radio Telescope (GMRT; \citealt{prasad-flagcal-2012}); and \textsc{serpent} for preprocessing data from the Multi-Element Radio Linked Interferometer Network (e-MERLIN; \citealt{serpent-peck-2013}). For RFI detection in MWA observations, LOFAR's \textsc{aoflagger} is used. This flagger has shown to have a good accuracy, is fast and has a library interface, which allows it to be integrated in a pipeline.

@@

\begin{table*}
\caption{List of observations used in the analyses. The RFI column contains the fraction of visibilities that the initial \textsc{aoflagger} run has classified as RFI (which includes false positives). The DVB column displays the fraction of visibilities that were unusable because of the presence of DVB signals, and which the flagger does not flag. Because the DVB frequencies are only observed in GLEAM observations, this fraction is only given for those. }\label{tbl:obs-list}
\begin{tabular}{|l|l|r|r|r|c|}
\hline
\textbf{Project} &\textbf{Date} & \textbf{Frequencies (MHz)} & \textbf{Duration} & RFI & DVB \\
\hline
EoR high& 2013-08-23 & 167.0--197.7              & 3h & 0.53\% & ---\\
GLEAM & 2013-08-25 & 72.3--133.8, 138.9--230.8 & 7h & 0.94\% & 0\% \\
EoR low& 2013-08-26 & 138.9--169.6              & 6h & 0.81\% & ---\\
GLEAM & 2013-11-05 & 72.3--133.8, 138.9--230.8 & 7h & 1.27\% & 0\% \\
GLEAM & 2013-11-25 & 72.3--133.8, 138.9--230.8 & 7h & 0.69\% & 0\% \\
EoR high& 2014-02-05 & 167.0--197.7              & 6h & 0.54\% & -- \\
GLEAM & 2014-03-16 & 72.3--133.8, 138.9--230.8 & 7h & 0.79\% & 0\% \\
GLEAM & 2014-03-17 & 72.3--133.8, 138.9--230.8 & 7h & 1.64\% & 3\% \\
EoR high& 2014-04-10 & 167.0--197.7              & 6h & 0.68\% & ---\\
GLEAM & 2014-06-18 & 72.3--133.8, 138.9--230.8 & 7h & 0.98\% & 0\% \\
RFI test&2014-08-27& 131.2--161.9     & 2$\times$8 m& 1.85\% & --- \\
\hline
\multicolumn{3}{|r|}{\textbf{Total}} & \textbf{63h} & \textbf{...\%} & \textbf{0.5\%}\\
\hline
\end{tabular}
\end{table*}

@@

%\noindent\begin{figure*}
%\begin{center}\hspace*{-0.2cm}\includegraphics[width=18cm]{img/plot-rfi-per-set-eor}\vspace{-1cm}
%\caption{..}
%\label{fig:rfi-per-set-eor}
%\end{center}
%\end{figure*}

%\noindent\begin{figure*}
%\begin{center}\hspace*{-0.2cm}\includegraphics[width=18cm]{img/plot-rfi-per-set-gleam}\vspace{-1cm}
%\caption{..}
%\label{fig:rfi-per-set-gleam}
%\end{center}
%\end{figure*}

@@

\noindent\begin{figure*}
\begin{center}\hspace*{-0.2cm}\includegraphics[width=18cm]{img/rfi-per-band}
\caption{RFI per subband}
\label{fig:rfi-per-band}
\end{center}
\end{figure*}

\noindent\begin{figure*}
\begin{center}\hspace*{-0.2cm}\includegraphics[width=18cm]{img/lognlogs-multiplot}
\caption{Log N log S}
\label{fig:lognlogs}
\end{center}
\end{figure*}
